Wolbachia, Toxoplasma and O. Cordyceps: Extreme Host-Manipulating Microbes That Changed the World

### 2015-12-16 paper.tex

\title{Wolbachia and Two Other Microbes Nobody Cares About That Changed The World Kinda}
    \author{C. Bekins, R. Chapman and B. Wasti}
    \date{December 2015}
    \maketitle

\section*{Introduction}
\section*{Wolbachia}
\textit{Wolbachia} is a group of bacteria that infects arthropod species. Technically there is only one true type of \textit{Wolbachia}, \textit{Wolbachia pipientis}, but due to the genetic diversity of different strains, it is commonly split into two groups, group A and group B, although it is generally just referred to as \textit{Wolbachia}.

@@

It goes back to the idea that there are two separate groups of \textit{Wolbachia}, group A and group B. Comparing the 16S rDNA does not give a full picture for the genetic diversity of the \textit{Wolbachia}, and the ftsZ gene was used to compare the genetics of 38 different \textit{Wolbachia} strains.\cite{Wgenetics} It was found that there is in fact a very large genetic difference between strains, and that group A and group B diverged 58-67 million years ago.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=.4\textwidth]{images/WolbachiaTree.jpg}
    \caption{ ... }
    \label{fig:wolbachiha_tree}
\end{figure}

>>>>

\title{Wolbachia, Toxoplasma and O. Cordyceps: Extreme Host-Manipulating Microbes That Changed the World}
    \author{C. Bekins, R. Chapman and B. Wasti}
    \date{December 2015}
    \maketitle

\section*{Introduction}
When thinking of infection most people will think of getting sick or of a swelling cut that is turning green. However, most people do not think about wandering around randomly and trying to climb a tree, only to have a fungus grow out of their spinal column. That is because humans are (mostly) lucky with respect to infections in this world; the worst we have to worry about is dying.

For other animals, dying is possibly the best case for an infection. There are a few microbes that will manipulate their host to extreme lengths. From hijacking the reproductive systems to complete mind control, and seemingly everything in between, the parasitic microbial world has developed a whole range of tools for which to survive at any cost. In this paper, we will go over three of these host-manipulating microbes, each with a different characteristic: \textit{Wolbachia}, the reproductive-system hijacker; \textit{Toxoplasma}, the subtle schizophrenic inducer; and the infamous \textit{O. Cordyceps}, the mind-controlling fungus.

\section*{Wolbachia}
\textit{Wolbachia} is a group of bacteria that infects arthropod species. Technically there is only one true type of \textit{Wolbachia}, \textit{Wolbachia pipientis}, but due to the genetic diversity of different strains, it is commonly split into two groups, group A and group B, although it is generally just referred to as \textit{Wolbachia}.

@@

It goes back to the idea that there are two separate groups of \textit{Wolbachia}, group A and group B. Comparing the 16S rDNA does not give a full picture for the genetic diversity of the \textit{Wolbachia}, and the ftsZ gene was used to compare the genetics of 38 different \textit{Wolbachia} strains.\cite{Wgenetics} It was found that there is in fact a very large genetic difference between strains, and that group A and group B diverged 58-67 million years ago. Figure \ref{fig:wolbachia_tree} shows a part of the tree for the strain distribution of Wolbachia.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=.4\textwidth]{images/WolbachiaTree.jpg}
    \caption{ A depiction of the strain distribution for Wolbachia. Most importantly, the separation between group A and group B is depicted. \cite{Wtree_image} }
    \label{fig:wolbachiha_tree}
\end{figure}
